An asymptotically normal out-of-sample test of equal predictive accuracy for nested models

### 2013-02-12 Paper.tex
Start corrections to bootstrap result for vcv convergence
The initial 'proof' of theorem 3--convergence of the bootstrap distribution
to the limiting distribution of the statistic--is pretty wrong.  I'm
fixing it by explicitly invoking de Jong and Davidson's vcv result for the
convergence of the bootstrap variance.  This commit adds part of it, but
the LLN for the deriviatives and stochastic equicontinuity still need to be
established.

\newcommand{\aic}{\textsc{aic}}
\newcommand{\bic}{\textsc{bic}}
\newcommand{\brc}{\textsc{brc}}
\newcommand{\cdf}{\textsc{cdf}}
\newcommand{\clt}{\textsc{clt}}
\newcommand{\dd}[1]{\frac{\partial}{\partial #1}}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\fclt}{\textsc{fclt}}
\newcommand{\fwe}{\textsc{fwe}}
\newcommand{\gdp}{\textsc{gdp}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\ma}{\textsc{ma}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\ned}{\textsc{ned}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\oos}{\textsc{oos}}
\newcommand{\sfwe}{\textsc{sfwe}}
\newcommand{\spa}{\textsc{spa}}
\newcommand{\wfwe}{\textsc{wfwe}}

@@

\author{Gray Calhoun\thanks{ Economics Department; Iowa State
    University; Ames, IA 50011.  Telephone: (515) 294-6271.  Email:
    \texttt{gcalhoun@iastate.edu}.  Web:
    \texttt{http://gray.clhn.co}.  I'd like to
    thank Helle Bunzel, Todd Clark, Graham Elliott, Yu-Chin Hsu,
    Michael McCracken, Pablo Pincheira, Allan Timmermann, Stephane
    Meng-Feng Yen and participants at the 2011 Midwest Econometrics
    Group meeting for helpful comments and discussions.  I'd also like to thank Amit
    Goyal for providing computer code and data for his 2008
    \textsc{rfs} paper with Ivo
    Welch \citep{GoW:08}.} \\
  Iowa State University}

@@

\begin{thm}\label{res:3}
  Suppose the following conditions hold:
  \begin{enumerate}
  \item Let $\{y_t, z_t\}$ be an $L_2$-\textsl{\ned} process of size $-\frac12$ on
    a strong mixing series $\{V_t\}$ of size $-\frac{r}{r-2}$, with $r
    > 2$.
  \item The sequence $(y_1^{*}, z_t^*),\dots,(y_T^{*}, z_T^*)$ is
    generated by the moving blocks, circular blocks, or stationary
    bootstrap with block lengths drawn from the geometric
    distribution, and $b$ satisfies $b \to \infty$ and $\frac{b}{P}
    \to 0$; $b$ is the block length of the moving or circular blocks
    bootstraps and is the expected block length of the stationary
    bootstrap.
  \item The estimator $\hat{\beta}_t$ of $\beta_0$ is estimated with a
    recursive, rolling, or fixed estimation window and satisfies
    $\hat{\beta}_{t} - \beta_{0} = \hat{B}_{t} H_t$; $\hat{B}_{t}$ is
    a sequence of $k$ by $q$ matrices such that $\sup_t |\hat{B}_t -
    B| \to^p 0$, $\sup_t |\hat{B}_t^{*} - B^{*}| \to^p 0$, and $B^{*}
    = B + o_p$; $H_{t}$ is a sequence of $q$-vectors such that 
    \begin{equation}
      H_{t} = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s} & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s} & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s} & \text{fixed window} \\
      \end{cases}
    \end{equation}
    $h_{s}(\beta) = h(y_{s}, z_{s}, \beta)$, $h_s = h_s(\beta_0)$, and
    $\E h_{s} = 0$ for all $s$; and $\frac12 \sum_{t=1}^T h_t^* = 0$
    a.s. where $h_t^* = h(y_s^*, z_s^*, \hat\beta_T)$.  Also define
    \begin{equation}
      H_{t}^* = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s}^* & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s}^* & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s}^* & \text{fixed window} \\
      \end{cases}
    \end{equation}
  \item Let $f_{t}(\beta) = f(y_{t}, z_{t}, \beta)$ be an $l$-vector;
    $\| \sup_{\beta \in N} f_{t}(\beta) \|_{r+\delta}$ and $\|
    \sup_{\beta \in N} h_{t}(\beta) \|_{r+\delta}$ are uniformly
    finite in $t$, where $N$ is an open neighborhood of $\beta_{0}$.
  \item Each element of $\E f_{t}(\beta)$ is continuously
    differentiable in the neighborhood $N$.  Moreover, $\sup_{\beta
      \in N} | F_{t}(\beta) |$ is uniformly finite in $t$ with
    $F_{t}(\beta) = \dd{\beta'} \E f_{t}(\beta)$.
  \item There exist finite constants $C$, $\phi > 0$, and $Q \geq r$
    such that \[\sup_{\epsilon : N(\beta_{0}, \epsilon) \subset N}\|
    \sup_{\beta \in N(\beta_{0}, \epsilon)} f_{t}(\beta) - f_{t} \|_Q
    \leq C \epsilon^{\phi}\] and \[\sup_{\epsilon : N(\beta_{0},
      \epsilon) \subset N}\| \sup_{\beta \in N(\beta_{0}, \epsilon)}
    h_{t}(\beta) - h_{t} \|_Q \leq C \epsilon^{\phi},\] where
    $N(\beta, \epsilon) = \{b : |b - \beta| < \epsilon\}$.
  \item $R, P \to \infty$ as $T \to \infty$.
  \item The asymptotic variance matrix of $\bar{f}(\beta_0)$ is
    uniformly positive definite and $\E f_t(\beta_0) = \E
    f_s(\beta_0)$ for all $s$ and $t$.
  \end{enumerate}
  Then
  \begin{equation}
    \pr[\sup_x | \pr^*[\sqrt{P} (\bar{f}^* - \bar{f}(\hat{\beta}_{T}))
        \leq x] - \pr[\sqrt{P}( \bar{f} - \E \bar{f}(\beta_0)) \leq x] | >
      \epsilon] \to 0
  \end{equation}
  for all $\epsilon > 0$.
\end{thm}

@@

\begin{rem}
  \citet{CoS:07} use the distribution of $\sqrt{P}(\bar{f}^{*} -
  \bar{f})$ to approximate that of $\sqrt{P}(\bar{f} - \E
  \bar{f}(\beta_0))$.  But it is clear that
  $\bar{f}(\hat{\beta}_{T+1})$ is the bootstrap analogue of $\E
  \bar{f}(\beta_0)$, the parameter of interest.  Because their
  bootstrap is miscentered, \citet{CoS:07} must redefine
  $\hat{\beta}_t^{*}$ to achieve consistency.  In this paper, though,
  consistency arises naturally.
\end{rem}

@@

\begin{lema}\label{res:a3}
  Suppose the conditions of Theorem~\ref{res:3} hold; let $\psi_{t} =
  (f_{t}' - \E f_{t}', a_t h_{t}')'$ and $\psi_t^{*} =
  (f_t^{*}(\hat{\beta}_{T+1})' - \bar{f}(\hat{\beta}_{T+1})',
  h_t^{*\prime})'$.  Then
  \begin{equation}\label{eq:1}
    \pr\Big[\sup_x \Big| \pr^{*}\Big[ \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \psi_{t}^{*}
    \leq x \Big] - \pr\Big[ \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \psi_{t}
    \leq x \Big] \Big| > \epsilon \Big] \to 0
  \end{equation}
  for all positive $\epsilon$.
\end{lema}

\begin{proof}[Proof of Lemma~\ref{res:a3}]
  Note that $\psi_t$ satisfies a \clt, \citep[Theorem~2]{Jon:97}; the
  bootstrap average obeys a \clt\ as well from
  \citet[Theorem~2]{Cal:11d} and is asymptotically normal, so it
  suffices to prove that
  \[ \tfrac{ 1 }{ P } \var^* \Big( \sum_{t=R+1}^T \psi_t^* \Big) -
  \tfrac{ 1 }{ P } \var \Big( \sum_{t=R+1}^T \psi_t \Big) \to^p 0. \]
  I will prove this result for the Moving Blocks Bootstrap---the proof
  for other block bootstraps is similar and follows as in
  \cite{Cal:11d}.

@@

  Note that
  \begin{equation*}
  \tfrac{ 1 }{ P } \var^* \Big( \sum_{t=R+1}^T \psi_t^* \Big)
  = \su \hat\psi_s \hat\psi_t'
     - \tilde \psi( \hat \beta_T ) \tilde \psi( \hat \beta_T )'
  \end{equation*}
  where
  \begin{equation*}
    \tilde \psi( \hat \beta_T
    = \tfrac{ 1 }{ T - b } \sum_{ t = b + 1 }^{ T - b } \psi_t( \hat \beta_T )
       + \tfrac{ 1 }{ T - b } \sum_{ t = 1 }^b \tfrac{ t }{ b }
         \big( \psi_t( \hat \beta_T ) + \psi_{ T - t + 1 }( \hat \beta_T ) \big)
  \end{equation*}
  Since $\bar \psi( \hat \beta_T ) = 0$ by construction and (finish explanation), we have
  \begin{align*}
    \tilde \psi( \hat \beta_T )
    &= \tfrac{ b }{ T } \tfrac{ 1 }{ T - b } \sum_{ t = b + 1 }^{ T - b } \psi_t( \hat \beta_T )
       + \tfrac{ 1 }{ T - b } \sum_{ t = 1 }^b 
       \Big( \tfrac{ t - b }{ b } - \tfrac{ b }{ T } \Big)
       \big( \psi_t( \hat \beta_T ) + \psi_{ T - t + 1 }( \hat \beta_T ) \big) \\
    &= o_p(1).
  \end{align*}
  Moreover,
  \begin{equation*}
    \su \hat\psi_s \hat\psi_t' \to^p \var \Big( \sum_{t=R+1}^T \psi_t \Big)
  \end{equation*}
  by (de Jong and Davidson reference), completing the proof.
\end{proof}

>>>>

\newcommand{\aic}{\textsc{aic}}
\newcommand{\bic}{\textsc{bic}}
\newcommand{\brc}{\textsc{brc}}
\newcommand{\cdf}{\textsc{cdf}}
\newcommand{\clt}{\textsc{clt}}
\newcommand{\dd}[1]{\frac{\partial}{\partial #1}}
\newcommand{\dgp}{\textsc{dgp}}
\newcommand{\fclt}{\textsc{fclt}}
\newcommand{\fwe}{\textsc{fwe}}
\newcommand{\gdp}{\textsc{gdp}}
\newcommand{\hac}{\textsc{hac}}
\newcommand{\lln}{\textsc{lln}}
\newcommand{\ma}{\textsc{ma}}
\newcommand{\mds}{\textsc{mds}}
\newcommand{\ned}{\textsc{ned}}
\newcommand{\ols}{\textsc{ols}}
\newcommand{\oos}{\textsc{oos}}
\newcommand{\sfwe}{\textsc{sfwe}}
\newcommand{\spa}{\textsc{spa}}
\newcommand{\wfwe}{\textsc{wfwe}}

@@

\author{Gray Calhoun\thanks{ Economics Department; Iowa State
    University; Ames, IA 50011.  Telephone: (515) 294-6271.  Email:
    \texttt{gcalhoun@iastate.edu}.  Web:
    \texttt{http://www.econ.iastate.edu/$\sim$gcalhoun}.  I'd like to
    thank Helle Bunzel, Todd Clark, Graham Elliott, Yu-Chin Hsu,
    Michael McCracken, Pablo Pincheira, Allan Timmermann, Stephane
    Meng-Feng Yen and participants at the 2011 Midwest Econometrics
    Group meeting for helpful comments and discussions.  I'd also like to thank Amit
    Goyal for providing computer code and data for his 2008
    \textsc{rfs} paper with Ivo
    Welch \citep{GoW:08}.} \\
  Iowa State University}

@@

\begin{thm}\label{res:3}
  Suppose the following conditions hold:
  \begin{enumerate}
  \item Let $\{y_t, z_t\}$ be an $L_2$-\textsl{\ned} process of size
    $-\frac12$ on the series $V_t$; $V_t$ is strong mixing of size
    $-\frac{r}{r-2}$ or uniform mixing of size $-\frac{r}{2r-2}$ with
    $r > 2$.\footnote{For uniform mixing processes, $r = 2$ is also
      allowed as long as $\sup_{\beta \in N} \psi_t(\beta)^2$ and
      $\sup_{\beta \in N} (\partial \psi_t(\beta) / \partial \beta)^2$
      are uniformly integrable; these terms are defined later in these
      assumptions.}  Moreover, the \ned-coefficients are uniformly
    bounded.
  \item The sequence $(y_1^{*}, z_t^*),\dots,(y_T^{*}, z_T^*)$ is
    generated by the moving blocks, circular blocks, or stationary
    bootstrap with block lengths drawn from the geometric
    distribution.  The (expected) block length $b$ satisfies $b \to
    \infty$ and $\frac{b}{P} \to 0$.
  \item The estimator $\hat{\beta}_t$ of $\beta_0$ is estimated with a
    recursive, rolling, or fixed estimation window and satisfies
    $\hat{\beta}_{t} - \beta_{0} = \hat{B}_{t} H_t$; $\hat{B}_{t}$ is
    a sequence of $k$ by $q$ matrices such that $\sup_t |\hat{B}_t -
    B| \to^p 0$, $\sup_t |\hat{B}_t^{*} - B^{*}| \to^p 0$, and $B^{*}
    = B + o_p$; $H_{t}$ is a sequence of $q$-vectors such that 
    \begin{equation}
      H_{t} = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s} & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s} & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s} & \text{fixed window} \\
      \end{cases}
    \end{equation}
    $h_{s}(\beta) = h(y_{s}, z_{s}, \beta)$, $h_s = h_s(\beta_0)$, and
    $\E h_{s} = 0$ for all $s$; and $\frac12 \sum_{t=1}^T h_t^* = 0$
    a.s. where $h_t^* = h(y_s^*, z_s^*, \hat\beta_T)$.  Also define
    \begin{equation}
      H_{t}^* = \begin{cases} 
        \tfrac1t \sum_{s=1}^t h_{s}^* & \text{recursive window} \\
        \tfrac1R \sum_{s=t-R+1}^t h_{s}^* & \text{rolling window} \\
        \tfrac1R \sum_{s=1}^R h_{s}^* & \text{fixed window} \\
      \end{cases}
    \end{equation}
  \item Let $f_{t}(\beta) = f(y_{t}, z_{t}, \beta)$ be an $l$-vector;
    the asymptotic variance matrix of $\bar{f}(\beta_0)$ is uniformly
    positive definite.
  \item Define $\psi_t(\beta) = (f_t(\beta), h_t(\beta))$;
    $\psi_t(\beta)$ is covariance stationary and is continuously
    differentiable in an open neighborhood $N$ of $\beta_0$.  Also,
    $\E( \sup_{\beta \in N} |\psi_t(\beta)|)^r$ and $\E\sup_{\beta \in
      N} |\frac{\partial \psi_t(\beta)}{\partial \beta}|^2$ are
    uniformly bounded.
  \item $R, P \to \infty$ as $T \to \infty$.
  \end{enumerate}
  Then
  \begin{equation}
    \pr[\sup_x | \pr^*[\sqrt{P} (\bar{f}^* - \bar{f}(\hat{\beta}_{T}))
        \leq x] - \pr[\sqrt{P}( \bar{f} - \E \bar{f}(\beta_0)) \leq x] | >
      \epsilon] \to 0
  \end{equation}
  for all $\epsilon > 0$.
\end{thm}

\begin{rem}
  \citet{Mcc:00} proves asymptotic normality under weaker smoothness
  condtions on $f_t(\beta)$ and $h_t(\beta)$: only their expectations
  must be continuously differentiable.  It may be possible to extend
  Theorem~\ref{res:3} to those weaker conditions, but the current
  proof relies on a theorem of \citepos{JoD:00} establishing the
  consistency of \hac\ estimators which requires differentiability of
  the observations.  Extending \citepos{JoD:00} result to
  nondifferentiable functions should be possible, but is beyond the
  scope of this paper.
\end{rem}

@@

\begin{rem}
  \citet{CoS:07} use the distribution of $\sqrt{P}(\bar{f}^{*} -
  \bar{f})$ to approximate that of $\sqrt{P}(\bar{f} - \E
  \bar{f}(\beta_0))$.  But it is clear that $\bar{f}(\hat{\beta}_T)$
  is the bootstrap analogue of $\E \bar{f}(\beta_0)$, the parameter of
  interest.  Because their bootstrap is miscentered, \citet{CoS:07}
  must redefine $\hat{\beta}_t^{*}$ to achieve consistency.  In this
  paper, though, consistency arises naturally.
\end{rem}

@@

\begin{lema}\label{res:a3}
  Suppose the conditions of Theorem~\ref{res:3} hold and let
  \[\psi_t^{*}(\beta) = 
  \binom{f_t^{*}(\beta) - \bar{f}(\beta)}{h_t^{*}(\beta) - \bar h(\beta)}.\]
  Then
  \begin{equation}\label{eq:1}
    \pr\Big[\sup_x \Big| \pr^{*}\Big[ \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \psi_{t}^{*}(\hat\beta_T)
    \leq x \Big] - \pr\Big[ \tfrac{1}{\sqrt{P}} \sum_{t=R+1}^T \psi_{t}(\beta_0)
    \leq x \Big] \Big| > \epsilon \Big] \to 0
  \end{equation}
  for all positive $\epsilon$.
\end{lema}

\begin{proof}[Proof of Lemma~\ref{res:a3}]
  Note that $\psi_t$ satisfies a \clt, \citep[Theorem~2]{Jon:97}; the
  bootstrap average obeys a \clt\ as well from
  \citet[Theorem~2]{Cal:11d}, so it suffices to prove that
  \[ \tfrac{ 1 }{ P } \var^* \Big( \sum_{t=R+1}^T \psi_t^*( \hat
  \beta_T ) \Big) - \tfrac{ 1 }{ P } \var \Big( \sum_{t=R+1}^T \psi_t(
  \beta_0 ) \Big) \to^p 0. \] I will prove this result for the Moving
  Blocks Bootstrap---the proof for other block bootstraps is similar
  and follows as in \cite{Cal:11d}.

@@

  Note that
  \begin{equation*}
  \tfrac{ 1 }{ P } \var^* \Big( \sum_{ t = R + 1 }^T \psi_t^*( \hat \beta_T ) \Big)
  = \su \hat \psi_s( \hat \beta_T ) \hat \psi_t( \hat \beta_T )'
     - \tilde \psi( \hat \beta_T ) \tilde \psi( \hat \beta_T )'
  \end{equation*}
  where
  \begin{equation*}
    \tilde \psi( \hat \beta_T )
    = \tfrac{ 1 }{ T - b } \sum_{ t = b + 1 }^{ T - b } \psi_t( \hat \beta_T )
       + \tfrac{ 1 }{ T - b } \sum_{ t = 1 }^b \tfrac{ t }{ b }
         \big( \psi_t( \hat \beta_T ) + \psi_{ T - t + 1 }( \hat \beta_T ) \big)
  \end{equation*}
  Since $\bar \psi( \hat \beta_T ) = 0$ by construction and (finish explanation), we have
  \begin{align*}
    \tilde \psi( \hat \beta_T )
    &= \tfrac{ b }{ T } \tfrac{ 1 }{ T - b } \sum_{ t = b + 1 }^{ T - b } \psi_t( \hat \beta_T )
       + \tfrac{ 1 }{ T - b } \sum_{ t = 1 }^b 
       \Big( \tfrac{ t - b }{ b } - \tfrac{ b }{ T } \Big)
       \big( \psi_t( \hat \beta_T ) + \psi_{ T - t + 1 }( \hat \beta_T ) \big) \\
    &= o_p(1).
  \end{align*}
  Moreover,
  \begin{equation*}
    \su \hat\psi_s \hat\psi_t' \to^p \var \Big( \sum_{t=R+1}^T \psi_t \Big)
  \end{equation*}
  by \citet[Theorem 2.2]{JoD:00}.  To see this, observe that 
  \begin{equation*}
    \su \hat\psi_s \hat\psi_t' 
    = \tfrac{1}{T} \sum_{s,t=1}^T \hat\psi_s \hat\psi_t' \;
      \big(1 - \tfrac{\lvert s - t \rvert}{b}\big)^+ + term.
  \end{equation*}
  The first term satisfies most of \citepos{JoD:00} Assumptions by
  construction and it suffices to prove their condition (2.9) holds,
  namely that
  \begin{equation}\label{eq:2}
    \sup_{\beta \in N} \Bigg\lVert \tfrac{1}{T} \sum_{t=1}^T e^{i \xi t/b} 
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} - 
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg) 
    \Bigg\rVert_2 \to 0
  \end{equation}
  for all $i$, $j$, and $\xi$.  I will show that \eqref{eq:2} holds
  for the real part of the expression; the proof for the complex part
  is similar.

  For any fixed $\beta \in N$,
  \begin{equation*}
    \tfrac{1}{T} \sum_{t=1}^T \cos( \xi t / b )
    \Bigg( \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} - 
          \E \frac{\partial \psi_{jt}(\beta)}{\partial \beta_i} \Bigg) 
  \end{equation*}
  obeys an \lln\ (give reference and explanation -- you need to show
  that the derivatives are \ned\ as well.) and convergence in $L_2$
  follows from uniform integrability.

  TODO: add stochastic equicontinuity conditions.
\end{proof}
